\subsection{New Functionality}
\subsubsection{Basic Functionality}

\begin{itemize}
\item Add a discretization text input option CRS for a real-world model coordinate reference system, for example, an EPSG integer code, authority string, or Open Geospatial Consortium Well-Known Text (WKT) specification. The entry for CRS does not affect the model simulation, but it is written to a version 2 binary grid file so that postprocessors can locate the grid in space.
\end{itemize}
\subsubsection{Model}

\begin{itemize}
\item Add DEPENDENT\_VARIABLE\_SCALING option to scale the dependent variable and the right-hand side to avoid very large positive or negative values in the GWT and GWE models. Specified outer and inner DVCLOSE values in the model IMS package will be relative closure criteria. This option applies the same scaling approach used in the MT3DMS GCG package.
\end{itemize}
\subsubsection{Stress Packages}

\begin{itemize}
\item Add HIGHEST\_SATURATED option to the GWT SRC package. When activated, the HIGHEST\_SATURATED option will apply a mass source loading rate to the specified CELLID or the highest underlying cell with a non-zero cell saturation. The HIGHEST\_SATURATED option has an additional complication for certain types of grids specified using the DISU Package. When the DISU Package is used, a cell may have more than one cell underlying it. If the overlying cell were to become inactive, there is no straightforward method for determining how to apportion the mass source loading rate to the underlying cells. In this case, the approach described by \cite{modflowusg} is used. The mass source loading rate is assigned to the first active cell encountered (determined by searching through the underlying cell numbers from the lowest number to the highest number). In this manner, the total mass source loading rate is conserved; however, the spatial distribution of the applied mass source loading rate may not be maintained as layers become dry or wet during a simulation.
\end{itemize}
\subsection{Bug Fixes And Other Changes To Existing Functionality}
\subsubsection{Internal Flow Packages}

\begin{itemize}
\item Fixed first-order decay in GWT so that production/decay does not occur when aqueous or sorbed concentration are negative.
\item Fixed bugs in the GWF VSC and BUY packages, affecting GWT simulations with multiple species that have different inactive regions.  In this uncommon situation, the concentrations used to calculate fluid density could erroneously be set to zero in areas with different inactive regions.
\end{itemize}
\subsubsection{Solution}

\begin{itemize}
\item With the PRT model's Particle Release Package's EXTEND\_TRACKING option enabled, particles beginning the simulation's final timestep in a stationary cell (an active cell with no flow across any boundary) were not terminated gracefully. Particles in these conditions will now terminate properly with status code 10 (timeout).
\end{itemize}
\subsubsection{Stress Packages}

\begin{itemize}
\item Fixed viscosity ratio assignment in the GWT MAW package when the GWF VSC package is active.
\end{itemize}
